---
title: "Data Science Project"
author: "Monica Chan"
date: "4/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

I decided to investigate NYC's education data, specifically the graduation rates in the five boroguhs of NYC: Manhattan, Queens, Brooklyn, The Bronx, and Staten Island. The original data set had data for these five boroughs from the years 2001-2006. There were a lot of other variables the set investigated, but I decided to focus on graduation rates in thise regions. This first chunk of code just sections off the data into the sets that I wanted to investigate, the individual graduation rates for each borough. The first table is the raw data I worked with, and the second table is the average graduation rate for each borough. I thought this was not the best measurement for better school systems, I thought instead that maybe investigating which borough's graduation rate improved the most on average would be a better measure for this. 

```{r echo = FALSE}
library(readr)
library(ggplot2)
BoroughEducation <- read.csv("BoroughEducation.csv")
class(BoroughEducation)
row1 = 1
row35 = 35
dataSet = BoroughEducation[row1:row35, 1:6]
class(dataSet)

#Data for individual boroughs
brooklyn <- dataSet[1:7,3:6]
bronx <- dataSet[8:14,3:6]
manhattan <- dataSet[15:21, 3:6]
queens <- dataSet[22:28,3:6]
statenIsland <- dataSet[29:35,3:6]

brookNumeric <- as.numeric(brooklyn[,4])
bronxNumeric <- as.numeric(bronx[,4])
manhattanNumeric <- as.numeric(manhattan[,4])
queensNumeric <- as.numeric(queens[,4])
statenNumeric <- as.numeric(statenIsland[,4]) 

print(dataSet)
avgGraduationRate <- c(mean(brookNumeric), mean(bronxNumeric), mean(manhattanNumeric), mean(queensNumeric), mean(statenNumeric))
names <- c("Brooklyn", "Bronx", "Manhattan", "Queens", "Staten Island")

averageGraduationRateNYC <-data.frame(names, avgGraduationRate)

print(averageGraduationRateNYC)
```

## Calculations
I wrote the function growthGrad to find the mean change in graduation rate from year to year. Essentially I just subtract year n from year n1... etc. and add that to a list. I then return the mean of that list. One thing I learned from this experience through some frsutrating debugging was that there are different types of vectors, and I needed to use a numeric vector instead of a list! 


```{r}
growthGrad <- function(x){
  previousYear <- x[1]
  years <- x
  z <- numeric()
  for(i in x[2:7]){
    y <- i - previousYear
    z <- c(z, y)
    previousYear = i
  }
  averageGrowth = mean(z)
  return(averageGrowth)
}
```

## Plots

It can be seen from this graph that the borough with the highest average graduation rate growth is Brooklyn! 

```{r echo = FALSE}
brooklynGrad = brooklyn[,4]
bronxGrad = bronx[,4]
manhattanGrad = manhattan[,4]
queensGrad = queens[,4]
statenIslandGrad = statenIsland[,4]

#data for individual questions
manhattanG <- growthGrad(manhattanGrad)
queensG <- growthGrad(queensGrad)
bronxG <- growthGrad(bronxGrad)
brooklynG <- growthGrad(brooklynGrad)
statenG<- growthGrad(statenIslandGrad)

gradGrowth <- c(manhattanG, queensG, bronxG, brooklynG, statenG)
boroughNames <- c("Manhattan", "Queens", "Bronx", "Brooklyn", "Staten Island")
gradData <- data.frame(boroughNames, gradGrowth)
```

```{r echo = FALSE}
g = ggplot(gradData, aes(x = boroughNames, y = gradGrowth)) + geom_bar(stat = "identity", fill="#6eadc4", colour="black")
g
```



